Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2021-01-07 manuscript/draft_rsfmri.Rmd

# load packages
library(car) # version 3.0.9
library(plyr) # version 1.8.6
library(dplyr) # version 1.0.2
library(tidyr) # version 1.1.2
library(kableExtra) # version 1.2.1
library(ggplot2) # version 3.3.2
library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
library(cowplot)
library(RColorBrewer) # version 1.1.2
library(psych) # version 2.0.8
#library(lsr)
#library(wesanderson) # colour palette
#library(haven)

library(lme4)
library(sjPlot)

library(knitr) # version 1.30
```

@@

# Results

## Quality control
We performed different quality controls for the rsfMRI data.
As described in @Beyer_2020, there was a significant interaction of group and time on mFD. In particular, the intervention group reduced in head motion in the follow-up assessments compared to the control group.
We also inspected DVARS, a measurement of whole-brain variation in signal which is often reflective of noise in the data. We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
```{r tableQA, fig.cap="Summary table of QA measures"}

for (i in 1:nrow(final)) {
  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
  conf = read.csv(
    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",
      tmp, "/confounds.csv"))
  tmp = cor.test(conf$DVARS, conf$FD)
  final[i, "corr_DVARS_FD"] = tmp$estimate
  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
  final
}
```

```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD'}
# ,  fig.align='center', }
# eval = TRUE, fig.asp = 0.7, fig.width = 6

@@

>>>>

# load packages
library(car) # version 3.0.9
library(plyr) # version 1.8.6
library(dplyr) # version 1.0.2
library(tidyr) # version 1.1.2
library(kableExtra) # version 1.2.1
library(ggplot2) # version 3.3.2
library(patchwork) # version 1.0.1; API for sequentially building up a plot (similar to gridExtra and cowplot)
library(cowplot)
library(RColorBrewer) # version 1.1.2
library(psych) # version 2.0.8

#library(lsr)
library(wesanderson) # colour palette
#library(haven)

library(lme4)

library(sjPlot) #only works for HTML output
library(texreg) #works for Rmarkdown
library(knitr) # version 1.30
```

@@

\newpage
# Results

## Quality control
We performed different quality controls for the rsfMRI data.
As described in @Beyer_2020, group and time show an interaction effect for mFD where the intervention group reduced in head motion in the follow-up assessments compared to the control group.
We also inspected DVARS, a measure of whole-brain signal variation which is reflective of noise in the data. We saw that DVARS did not qualitatively differ between groups and time points, and that the correlation of DVARS and mFD remained similar over time points.
This indicates that even though there was a decrease in head motion in the intervention versus control group, the impact of head motion on imaging data remained similar over the study period.
```{r tableQcfc, echo=FALSE}
res <- read.csv("./report/FDFCcorrelations.csv")
rownames(res) <- c("minimally processed", "AROMA", "AROMA + CC", "AROMA + CC + GSR")
colnames(res) <- c('mean FD-QC', 'median FD-QC', 'sig. vertex', 'sig. vertex BH', 'distance-FD-QC', 'pval')

tab_qc_fc <- mk_tableQcfc(res)
tab_qc_fc
```
Further, we evaluated the impact of our denoising pipelines on different quality metrics following [@Ciric_2018]. Table \@ref(tab:tableQcfc) shows that in the minimally preprocessed data, mFD and FC between nodes was significantly positively correlated (mean Spearman's r = `r round(res[1,1],2)`). This relationship was negatively associated with distance between nodes (mean Spearman's r = `r round(res[1,5],2)`).  
Denoising procedures (AROMA, AROMA+CC, AROMA+CC+GSR) gradually reduced the correlation of mFD and connectivity, and reduced the distance dependency. Contrary to previous reports, GSR did not exercabate mFD-FC distance dependency. Overall, AROMA+CC+GSR performed best, with almost no mFD-FC correlation (mean Spearman's r = `r round(res[4,1],2)`), and minimal positive distance dependence (mean Spearman's r = `r round(res[4,5],2)`). Yet, AROMA+CC also performed well, and we thus calculated our analysis models for both preprocessing schemes.

```{r calculate summary of QA, eval=FALSE, echo=FALSE}
for (i in 1:nrow(final)) {
  tmp = strsplit(toString(final[i, "scan_dir"]), '/')[[1]][8]
  conf = read.csv(
    paste0("/data/pt_02161/preprocessed/resting/detailedQA/metrics/",
      tmp, "/confounds.csv"))
  tmp = cor.test(conf$DVARS, conf$FD)
  final[i, "corr_DVARS_FD"] = tmp$estimate
  final[i, "mean_DVARS"] = mean(conf$DVARS, na.rm = TRUE)
  final[i, "max_DVARS"] = max(conf$DVARS, na.rm = TRUE)
  final
}
```

```{r figDvarsmFD, fig.align = "center", out.width="100%", fig.cap='Summary of quality measures DVARS and mFD', echo=FALSE}
final=read.csv(file = "./report/all_QA_measures.csv")

@@

<!-- Neurovault -->
Unthresholded maps for the t-tests of the DMN and reward network (which were used for data aggregation) as well as contrasts of group, time and group-by-time interaction are on [NeuroVault](insert link here).
